Reinforcement Learning Report

### 2022-08-08 report/rl_qf_report.tex
make comparison for three algorithms

\section{Actor-Critic, Reinforce, and PPO}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms:
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{iteration \(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where 
    \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
    \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
    \EndFor
    \Loop \hspace{0.3mm} (update multiple times)
    \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \min{\left(
            \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
            {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}A_t, g(\epsilon, A_t)
    \right)}\)
    \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
    \EndLoop
\EndFor
\end{algorithmic}
\end{algorithm}

>>>>

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms:
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{iteration \(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where 
    \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
    \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
    \EndFor
    \Loop \hspace{0.3mm} (update multiple times)
    \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \min{\left(
            \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
            {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}A_t,
            \text{clip}\left(
                \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
                {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)},
                1-\epsilon, 1+\epsilon
            \right)A_t
    \right)}\)
    \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
    \EndLoop
\EndFor
\end{algorithmic}
\end{algorithm}

\subsection{Training result}
To make better comparison, the policy function of the two algorithms
is implemented the same way as we did in Actor-Critic algorithm, and the same 
is true for the value function used in PPO method. 
\vspace{3mm}\newline The following graph shows the training result of these 
three algorithms (300 moving average reward vs iteration), with 600 iterations
(5 episodes per iteration), \(T=300\), \(\gamma=0.99\), 
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{center}
    \includegraphics[scale=0.67]{out}
\end{center}
As is shown in the graph, the Actor-Critic algorithm converges the fastest,
within 50 iterations, followed by Reinforce algorithm, which converges within
about 250 iterations, despite a small drop of moving average around iteration 
300, which is quickly corrected by the algorithm within 50 iterations. PPO 
algorithm converges the slowest, with about 300 iterations. Despite of the 
speed of convergence, the moving average rewards of convergence for these 
three algorithms are nearly the same (around 400), this is as expected.
